---
title: "Homework 6"
author: "Lisa Eisler"
date: "11/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "90%"
)

library(tidyverse)
library(knitr)
library(readxl)
library(haven)
library(plotly)
library(viridis)
library(broom)
library(purrr)
library(modelr)
library(mgcv)
library(patchwork)

set.seed(1)
```

## Problem 1

Read in and clean birthweight data for regression.

```{r, message=FALSE}
birthweight_data = read_csv(file = "./data/birthweight.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    babysex = factor(babysex, levels =  c("1", "2"), labels = c("male", "female")),
    frace = factor(frace, levels =  c("1", "2", "3", "4", "8", "9"), labels = c("white", "black", "asian", "puerto rican", "other", "unknown")),
    malform = factor(malform, levels =  c("0", "1"), labels = c("absent", "present")),
    mrace = factor(mrace, levels =  c("1", "2", "3", "4", "8"), labels = c("white", "black", "asian", "puerto rican", "other"))
  )
birthweight_data
```

```{r}
skimr::skim(birthweight_data)
```
*There are no missing data*

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process. 

*Based on my clinical work in the delivery room and NICU, I propose that birthweight is predicted by: gestationl age, length, head circumference, smoking status, income, and the number of prior low birthweight babies born to the mother.*

```{r} 
bwtmodel_1 = lm(bwt ~ gaweeks + bhead + blength + smoken + fincome + pnumlbw, data = birthweight_data)
bwtmodel_1 %>%
  summary()
```

*show table with parameters and p-value*

```{r}
bwtmodel_1 %>%
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 5)
```

Show a plot of model residuals against fitted values.

```{r} 
birthweight_data %>%
modelr::add_residuals(bwtmodel_1) %>%
modelr::add_predictions(bwtmodel_1) %>% 
ggplot(aes(x = pred, y = resid)) + geom_point(alpha = 0.5) + 
           geom_hline(yintercept = 0, color = "blue") + 
  labs(
    title = "Plot of model residuals against fitted values",
    x = "Predicted Birtweight (grams)",
    y = "Residuals"
   )
```

Create comparison models: One using length at birth and gestational age as predictors (main effects only) and one using head circumference, length, sex, and all interactions (including the three-way interaction) between these.

```{r, message=FALSE} 
bwtmodel_2 = lm(bwt ~ gaweeks + blength, data = birthweight_data) 
bwtmodel_3 = lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = birthweight_data)
```

Make a comparison of these two models with your model in terms of the cross-validated prediction error.

```{r} 
cv_bwtmodels =
  crossv_mc(birthweight_data, 100) %>%
    mutate( 
      train = map(train, as_tibble),
      test = map(test, as_tibble)
    )

cv_bwtmodels = 
  cv_bwtmodels %>%
    mutate(
      bwtmodel_1  = map(train, ~lm(bwt ~ gaweeks + bhead + blength + smoken + fincome + pnumlbw, data = .x)),
      bwtmodel_2  = map(train, ~lm(bwt ~ gaweeks + blength, data = .x)),
      bwtmodel_3  = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = as_tibble(.x)))) %>% 
    mutate(
      rmse_bwtmodel_1 = map2_dbl(bwtmodel_1, test, ~rmse(model = .x, data = .y)),
      rmse_bwtmodel_2   = map2_dbl(bwtmodel_2, test, ~rmse(model = .x, data = .y)),
      rmse_regmodel_3 = map2_dbl(bwtmodel_3, test, ~rmse(model = .x, data = .y)))

cv_bwtmodels %>% 
  select(starts_with("rmse")) %>% 
pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse,  fill = model)) + geom_violin(alpha = 0.5) 
```

*Based on RMSE, the model I developed (bwtmodel_1) appears to be the best fit for the data, and in addition contains a moderate amount of complexity compared with models 2 (too simple) and 3 (too complex with interaction terms).*

## Problem 2 

Download Central Park Weather Data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

Focus on a simple linear regression with tmax as the response and tmin as the predictor:

```{r}
lm(tmax ~ tmin, data = weather_df) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

We are interested in the distribution of two quantities estimated from these data: r^2 and log(β^0∗β^1). Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities.

```{r}
weather_estimates =
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy), 
    glance = map(models, broom::glance)) %>% 
  select(results, glance) %>% 
  unnest(results, glance) %>% 
  janitor::clean_names() %>% 
  select(term, estimate, r_squared) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate)  %>% 
  janitor::clean_names() %>% 
  mutate(
    log_betas = log(intercept * tmin)
    ) 
```

Plot the distribution of your estimates.

```{r}
rsquared_plot = 
  weather_estimates %>% 
  ggplot(aes(x = r_squared)) +
  geom_density() +
  labs(
    title = "Estimation Distributions",
    x = "r-squared",
    y = "Density"
  )

log_betas_plot = 
  weather_estimates %>% 
  ggplot(aes(x = log_betas)) +
  geom_density() +
   labs(
    x = "log(β0*β1)",
    y = "Density"
  )

(rsquared_plot + log_betas_plot)
```
*These plots demonstrate that the estimates are approximately normally distributed.*

Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(β^0∗β^1).

```{r}
pull(weather_estimates, r_squared) %>% 
  quantile(c(0.025, 0.975))

pull(weather_estimates, log_betas) %>%
  quantile(c(0.025, 0.975))
```

*The 95% CI for r-squared is (0.89, 0.93). The 95% CI for log(β^0∗β^1) is (1.97, 2.06).* 
