---
title: "Homework 6"
author: "Lisa Eisler"
date: "11/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(readxl)
library(haven)
library(plotly)
library(viridis)
library(broom)
library(purrr)
```

## Problem 1

Read in and clean birthweight data for regression. Create dichotomous maternal race variable for regression (black vs. non-black/unknown).

```{r, message=FALSE}
birthweight_data = read_csv(file = "./data/birthweight.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    babysex = factor(babysex, levels =  c("1", "2"), labels = c("male", "female")),
    frace = factor(frace, levels =  c("1", "2", "3", "4", "8", "9"), labels = c("white", "black", "asian", "puerto rican", "other", "unknown")),
    malform = factor(malform, levels =  c("0", "1"), labels = c("absent", "present")),
    mrace = factor(mrace, levels =  c("1", "2", "3", "4", "8"), labels = c("white", "black", "asian", "puerto rican", "other"))
  ) %>%
  mutate(mblack = ifelse(mrace == "black", "true", "false"))
birthweight_data
```

```{r}
sum(!complete.cases(birthweight_data))
```

*There are no missing data*

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

*Based on my clinical in the delivery room and NICU, I propose that birthweight is modeled by the following factors: gestationl age, length, head circumference, smoking status, income, and the number of prior low birthweight babies born to the mother. I will also include maternal race (black/non-black) given recent NYtimes coverage of the healthcare disparities impacting black mothers and their babies even after other factors such as SES are accounted for*

```{r, message=FALSE} 
linearmodel_1 = lm(bwt ~ gaweeks + bhead + blength + smoken + fincome + pnumlbw, data = birthweight_data) 
linearmodel_1 %>% 
broom::tidy() %>% 
  select(term, estimate, p.value) 
```

```{r, message=FALSE} 
birthweight_data %>%
modelr::add_residuals(linearmodel_1) %>%
modelr::add_predictions(linearmodel_1) %>% 
ggplot(aes(x = pred, y = resid)) + geom_point(alpha = 0.5) + 
           geom_hline(yintercept = 0, color = "green") + 
  labs(
    title = "Plot of model residuals against fitted values",
    x = "Predicted Birtweight (grams)",
    y = "Residuals"
   )
```